---
title: "lab5"
author: "Claire Madden"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

## Attach packages

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)


```

## Monthly US energy consumption (renewables)

```{r}
# read in data, clean column names
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

```

- Make descriptions all lowercase
- only keep observations where "description" variable contains "consumption"
- remove any observations where "description" variable contains "total"

```{r}

renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))


```

### Convert 'yyyymm' column to a date

- parse current column with lubridate
- convert to tsibble friendly fomat so we can use with feast and fable
- convert value column to numeric (was read in as character)
- drop any NA in either month_sep or value

```{r}

# lubridate autofills and nonsensical dates with NA, how great!

renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# make a version where month and year are in separate columns to use later

renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))


```

### Check it out!

```{r}

renew_gg <- ggplot(data = renew_date, aes(x = month_sep, 
                                          y = value,
                                          group = description))+
  geom_line(aes(color = description))

renew_gg # saved graph with this name, can keep building by just referencing this name

```


Updating colors with paletteer palettes:
```{r}
renew_gg + 
  scale_color_paletteer_d("nationalparkcolors::CraterLake")
  


```

### Coerce renew_parsed to a tsibble (timeseries enabled df)
- usually modeling and forecasting in feast and forecast works better if you do this first

```{r}
# key: specify variable to be primary grouping 
# index: tsibble compatible time variable in df
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)

```

Let's look at our timeseries data in a few different ways:

```{r}
# this is the same as the ggplot we created before, done a different way
renew_ts %>% autoplot(value)

# breaks data into each "description" variable, by "month", across years
renew_ts %>% gg_subseries(value)

# look at season plot, within each season plot each year separately to see how things change
#renew_ts %>% gg_season(value) # we expected this to break...

# going to make the same graph with ggplot
ggplot(data = renew_parsed, aes(x = month, y = value, group = year))+
  geom_line(aes(color = year))+
  facet_wrap(~description, 
             ncol = 1, 
             scales = "free", 
             strip.position = "right") # description names on right
```

### Just look at hydroelectric energy consumption

```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")


hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
#hydro_ts %>% gg_season(value) still not workin

ggplot(hydro_ts, aes(x = month, y = value, group = year))+
  geom_line(aes(color = year))
```

### Calculate quarterly average consumption for hydropower
- index by (instead of group_by, specific for tsibble data)

```{r}

hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>%  # "." means "based on different groups that already exist"
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)
```

### Decompose the hydro_ts df

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% autoplot()
```

```{r}
# check distribution of remainder values, aiming for close to normal distribution
hist(components(dcmp)$remainder)
```



Now let's look at the ACF:
```{r}

hydro_ts %>% 
  ACF(value) %>% 
  autoplot()

# seasonality indicated in autocorrleation, observations that are 12 months apart are more highly correlated than observations that are any other distance apart

```


# DANGER DANGER - modeling and forecasting

```{r}

hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value),
    ETS(value) # can add a second model to show different forecast outcomes
  ) %>% 
  fabletools::forecast(h = "4 years") #how long into the future do you want the forecast 

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep)>2010))


```

### Make a world map!


```{r}

world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"), 
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")


# mapview is a good quick way to view spatial data

mapview(world)

```











